---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/sfweight)](https://CRAN.R-project.org/package=sfweight)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->


sfweight is an opinionated translation of the wonderful spdep package. The goal is to provide a streamlined method of doing spatial statistics that works with sf objects, data frames, and the tidyverse. spdep is more flexible with the types of input objects and a bit more idiosyncratic in the syntax that is used. 

## Installation

You can install the development version from GitHub with 
```{r eval=FALSE}
remotes::install_github("Josiahparry/sfweight")
```


## Motivating examples

### Spatial OLS

We can fit a spatial Durbin model by calculating spatially lagged predictors. 

```{r, warning = FALSE, message=FALSE}
library(sfweight)
library(tidyverse)

acs_lagged <- acs %>% 
  mutate(nb = st_neighbors(geometry),
         wts = st_weights(nb),
         trans_lag = st_lag(by_pub_trans, nb, wts),
         bach_lag = st_lag(bach, nb, wts))


durbin_lm <- lm(med_house_income ~ trans_lag + by_pub_trans + bach_lag + bach, 
   data = acs_lagged)

broom::tidy(durbin_lm)
```

### Local Autocorrelation

We can create a Moran plot by creating a spatially lagged variable. Additionally the function `categorize_lisa()` will categorize high-high, high-low, etc., groupings of these variables.

```{r}
acs_lagged %>% 
  mutate(inc_lag = st_lag(med_house_income, nb, wts),
         lisa_group = categorize_lisa(med_house_income, inc_lag)) %>% 
  ggplot(aes(med_house_income, inc_lag, color = lisa_group)) +
  geom_vline(aes(xintercept = mean(med_house_income)), lty = 2, alpha = 1/3) +
  geom_hline(aes(yintercept = mean(inc_lag)), lty = 2, alpha = 1/3) + 
  geom_point() +
  labs(title = "Moran Plot",
       y = "Med. HH Income Spatial Lag",
       x = "Median Household Income") +
  theme_minimal() +
  scale_x_continuous(labels = scales::dollar) + 
  scale_y_continuous(labels = scales::dollar)
```


We can also calculate the Local  Moran's I for each observation using the function `local_moran()` this will create a dataframe column containing the I, expected I, variance, Z-value, and P-value for each observation. You can extract this using `tidyr::unpack()`. In order to do so you need to cast as a tibble then cast back to an sf object if you want to maintain the sf class. 


```{r}
acs_lisa <- acs_lagged %>% 
  mutate(lisa = local_moran(bach, nb, wts)) %>% 
  as_tibble() %>% 
  unpack(lisa) 

acs_lisa %>% 
  select(last_col(4:0))
```



```{r}
library(sf)
 
acs_lisa %>% 
  st_as_sf() %>% 
  ggplot(aes(fill = ii)) + 
  geom_sf(color = "black", lwd = 0.2) +
  scale_fill_binned(n.breaks = 5) + 
  theme_minimal()
```

## Basic usage & contiguities

```{r}
str(acs)
```

We can get neighbors based on Queen contiguities with `st_neighbors()`.


```{r}
nbs <- st_neighbors(acs)

nbs[1:5]
```

If needed, we can also identify the cardinalities from the neighbors list as well. 

```{r}
st_cardinalties(nbs)
```


We can get the weights from the neighbor contiguities as well. By default, `st_weights()` uses row standardization. 

```{r}
wts <- st_weights(nbs)

wts[1:5]
```

We can also calculate the spatial lag with the weights and neighbors. 

```{r}
inc_lag <- st_lag(acs$med_house_income, nbs, wts)

inc_lag[1:5]
```


## K-Nearest Neighbor Distances

If we have point data we can also identify the k-nearest neighbors with `st_knn()`. For an example we can use the `airbnb` dataset that's imported with `sfweight`. 

```{r}
airbnb
```

```{r message=FALSE, warning=FALSE}
airbnb_knn <- st_knn(airbnb)

airbnb_knn[1:5]
```

## Other weights

Point based weights implemented based on Luc Anselin and Grant Morrison's [notes](https://spatialanalysis.github.io/lab_tutorials/Spatial_Weights_as_Distance_Functions.html#kernal-weights). 

Inverse distance band

```{r}
airbnb_idw <- st_inverse_weights(airbnb$geometry, airbnb_knn)

airbnb_idw[1]
```

### Kernel based weights


Available kernels are:

- uniform
- triangular
- epanechnikov
- quartic
- gaussian

```{r}
airbnb_gauss <- st_kernel_weight(airbnb$geometry, airbnb_knn, "gaussian")

airbnb_gauss[1]
```


## Higher order neighbors 

```{r}
acs %>% 
  transmute(nb = st_neighbors(geometry),
            nb_2 = st_neighbor_lag(nb, 2),
            nb_cumul_2 = st_neighbor_lag_cumul(nb, 2))
```


