---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sfweight

<!-- badges: start -->
<!-- badges: end -->

The goal of sfweight is to create a tidier and more streamlined interface to the spdep package. The spdep package has an idiosyncratic syntax that can be difficult to fit into a typical data science workflow. sfweight creates a simpler interface to the spdep package. 

The intention behind sfweight is implement a simpler, but stricter workflow that enables the creation of neighbors, spatial weights, and spatially lagged variables. This will be accomplished by decoupling neighbors from weights and utilizing list objects.

sfweight uses sf objects whereas spdep is more flexible with the types of input objects vaialable. 

## Installation

You can install the development version from GitHub with 
```{r eval=FALSE}
remotes::install_github("Josiahparry/sfweight")
```


## Motivating examples

### Spatial OLS

We can fit a spatial Durbin model by calculating spatially lagged predictors. 

```{r, warning = FALSE, message=FALSE}
library(sfweight)
library(tidyverse)

acs_lagged <- acs %>% 
  mutate(nb = st_neighbors(geometry),
         wts = st_weights(nb),
         trans_lag = st_lag(med_house_income, nb, wts),
         bach_lag = st_lag(bach, nb, wts))


durbin_lm <- lm(med_house_income ~ trans_lag + by_pub_trans + bach_lag + bach, 
   data = acs_lagged)

broom::tidy(durbin_lm)
```

### Local Autocorrelation

We can create a Moran plot by creating a spatially lagged variable. Additionally the function `categorize_lisa()` will categorize high-high, high-low, etc., groupings of these variables.

```{r}
acs_lagged %>% 
  mutate(inc_lag = st_lag(med_house_income, nb, wts),
         lisa_group = categorize_lisa(med_house_income, inc_lag)) %>% 
  ggplot(aes(med_house_income, inc_lag, color = lisa_group)) +
  geom_vline(aes(xintercept = mean(med_house_income)), lty = 2, alpha = 1/3) +
  geom_hline(aes(yintercept = mean(inc_lag)), lty = 2, alpha = 1/3) + 
  geom_point() +
  labs(title = "Moran Plot",
       y = "Med. HH Income Spatial Lag",
       x = "Median Household Income") +
  theme_minimal() +
  scale_x_continuous(labels = scales::dollar) + 
  scale_y_continuous(labels = scales::dollar)
```


We can also calculate the Local  Moran's I for each observation using the function `local_moran()` this will create a dataframe column containing the I, expected I, variance, Z-value, and P-value for each observation. You can extract this using `tidyr::unpack()`. In order to do so you need to cast as a tibble then cast back to an sf object if you want to maintain the sf class. 


```{r}
acs_lisa <- acs_lagged %>% 
  mutate(lisa = local_moran(bach, nb, wts)) %>% 
  as_tibble() %>% 
  unpack(lisa) 

acs_lisa %>% 
  select(last_col(4:0))
```



```{r}
library(sf)
 
acs_lisa %>% 
  st_as_sf() %>% 
  ggplot(aes(fill = ii)) + 
  geom_sf(color = "black", lwd = 0.2) +
  scale_fill_binned(n.breaks = 5) + 
  theme_minimal()
```

## Basic usage & contiguities

```{r}
str(acs)
```

We can get neighbors based on Queen contiguities with `st_neighbors()`.


```{r}
nbs <- st_neighbors(acs)

nbs[1:5]
```

If needed, we can also identify the cardinalities from the neighbors list as well. 

```{r}
st_cardinalties(nbs)
```


We can get the weights from the neighbor contiguities as well. By default, `st_weights()` uses row standardization. 

```{r}
wts <- st_weights(nbs)

wts[1:5]
```

We can also calculate the spatial lag with the weights and neighbors. 

```{r}
inc_lag <- st_lag(acs$med_house_income, nbs, wts)

inc_lag[1:5]
```


## K-Nearest Neighbor Distances

If we have point data we can also identify the k-nearest neighbors with `st_knn()`. For an example we can use the `airbnb` dataset that's imported with `sfweight`. 

```{r}
airbnb
```

```{r message=FALSE, warning=FALSE}
airbnb_knn <- st_knn(airbnb)

airbnb_knn[1:5]
```

## Other weights

Point based weights implemented based on Luc Anselin and Grant Morrison's [notes](https://spatialanalysis.github.io/lab_tutorials/Spatial_Weights_as_Distance_Functions.html#kernal-weights). 

Inverse distance band

```{r}
airbnb_idw <- st_inverse_weights(airbnb$geometry, airbnb_knn)

airbnb_idw[1]
```

### Kernel based weights


Available kernels are:

- uniform
- triangular
- epanechnikov
- quartic
- gaussian

```{r}
airbnb_gauss <- st_kernel_weight(airbnb$geometry, airbnb_knn, "gaussian")

airbnb_gauss[1]
```


## Higher order neighbors 

```{r}
acs %>% 
  transmute(nb = st_neighbors(geometry),
            nb_2 = st_neighbor_lag(nb, 2),
            nb_cumul_2 = st_neighbor_lag_cumul(nb, 2))
```


